Pipeline Performance Figure Generation

### 2023-04-06 TestingPipeline/Scripts/paper_figure_generation.Rmd
Updating SI figures.

for (i in c(0,1)){
  for (j in c(0,1)){
    for (k in c(0,1)){
      for (l in c(0,1)){
        for (m in c(0,1)){
          for (n in c(0,1)){
            combos_list$toolcombo[p] <- paste(i,j,k,l,m,n)
            combos_list$toolcombo2[p] <- paste(if(i){"tv"}else{"0"},if(j){"DVF"}else{"0"},
                                               if(k){"tnv"}else{"0"},if(l){"VB"}else{"0"},
                                               if(m){"VS"}else{"0"},if(n){"VS2"}else{"0"})
            combos_list$tune_not_viral[p] <- i
            combos_list$DVF[p] <- j
            combos_list$tune_viral[p] <- k
            combos_list$VIBRANT[p] <- l
            combos_list$VS[p] <- m
            combos_list$VS2[p] <- n
            p <- p+1
          }
        }
      }
    }
  }
}

@@

ggplot(accuracy_scores, aes(x=toolcombo, y=precision, 
                                  color=numrules, fill=numrules)) +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
  ylab("Precision") +
  scale_fill_manual(name="Number of Rules",
                     values = alpha(rev(pal(6)), 0.5)) +
  scale_color_manual(name="Number of Rules",
                     values = alpha(rev(pal(6)), 1))
```

@@

ggplot(accuracy_scores, aes(x=toolcombo, y=recall, 
                                  color=numrules, fill=numrules)) +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
  ylab("Recall") +
  scale_fill_manual(name="Number of Rules",
                     values = alpha(rev(pal(6)), 0.5)) +
  scale_color_manual(name="Number of Rules",
                     values = alpha(rev(pal(6)), 1))
```

@@

precision vs recall
```{r}
ggplot(accuracy_scores, aes(x=precision, y=recall, 
                                  color=numrules, fill=numrules)) +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=20),
  ) +
  xlab("Precision") +
  ylab("Recall") +
  scale_fill_manual(name="Number of Rules",
                     values = alpha(rev(pal(6)), 0.5)) +
  scale_color_manual(name="Number of Rules",
                     values = alpha(rev(pal(6)), 1))

@@

making heatmap for the x-axis label
```{r}
ann_col = data.frame(
    tnv = combos_list$tune_viral,
    tv = combos_list$tune_not_viral,
    dvf = combos_list$DVF,
    vb = combos_list$VIBRANT,
    vs = combos_list$VS,
    vs2 = combos_list$VS2
)
rownames(ann_col) <- combos_list$toolcombo

@@

>>>>

for (i in c(0,1)){
  for (j in c(0,1)){
    for (k in c(0,1)){
      for (l in c(0,1)){
        for (m in c(0,1)){
          for (n in c(0,1)){
            combos_list$toolcombo[p] <- paste(i,j,k,l,m,n)
            combos_list$toolcombo2[p] <- paste(if(i){"tv"}else{"0"},if(j){"DVF"}else{"0"},
                                               if(k){"tnv"}else{"0"},if(l){"VB"}else{"0"},
                                               if(m){"VS"}else{"0"},if(n){"VS2"}else{"0"})
            combos_list$tune_not_viral[p] <- k
            combos_list$DVF[p] <- j
            combos_list$tune_viral[p] <- i
            combos_list$VIBRANT[p] <- l
            combos_list$VS[p] <- m
            combos_list$VS2[p] <- n
            p <- p+1
          }
        }
      }
    }
  }
}

@@

ggplot(accuracy_scores, aes(x=toolcombo, y=precision, 
                                  color=numrules, fill=numrules)) +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
  ylab("Precision") +
  scale_fill_manual(name="Number of Rules",
                     values = alpha(rev(viridis(6)), 0.5)) +
  scale_color_manual(name="Number of Rules",
                     values = alpha(rev(viridis(6)), 1))
```

@@

ggplot(accuracy_scores, aes(x=toolcombo, y=recall, 
                                  color=numrules, fill=numrules)) +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
  ylab("Recall") +
  scale_fill_manual(name="Number of Rules",
                     values = alpha(rev(viridis(6)), 0.5)) +
  scale_color_manual(name="Number of Rules",
                     values = alpha(rev(viridis(6)), 1))
```

@@

precision vs recall
```{r}
ggplot(accuracy_scores, aes(x=precision, y=recall, 
                                  color=numrules, fill=numrules)) +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=20),
  ) +
  xlab("Precision") +
  ylab("Recall") +
  scale_fill_manual(name="Number of Rules",
                     values = alpha(rev(viridis(6)), 0.5)) +
  scale_color_manual(name="Number of Rules",
                     values = alpha(rev(viridis(6)), 1))

@@

making heatmap for the x-axis label
```{r}
ann_col = data.frame(
    tnv = combos_list$tune_not_viral,
    tv = combos_list$tune_viral,
    dvf = combos_list$DVF,
    vb = combos_list$VIBRANT,
    vs = combos_list$VS,
    vs2 = combos_list$VS2
)
rownames(ann_col) <- combos_list$toolcombo

@@

# numbers in common between two tool combos
```{r}
combos_list_two <- combos_list
combos_list_two$numtools <- str_count(combos_list_two$toolcombo, "1")
combos_list_two <- combos_list_two[combos_list_two$numtools==1,]

two_tool_pairs <- num_high_viruses[rownames(num_high_viruses) %in% combos_list_two$toolcombo,
                                   colnames(num_high_viruses) %in% combos_list_two$toolcombo]

colnames(two_tool_pairs) <- c("VS2", "VS", "VB", "tv", "DVF", "tnv")
rownames(two_tool_pairs) <- c("VS2", "VS", "VB", "tv", "DVF", "tnv")
```

```{r}
viral_scores_two <- viral_scores[,colnames(num_high_viruses) %in% combos_list_two$toolcombo]

n <- 6
viral_scores_two_comb <- matrix(data=NA, nrow=n, ncol=n)

for (i in 1:n) {
  for (j in 1:n) {
    if (i==j) {
      viral_scores_two_comb[i,j] <- sum(viral_scores_two[,i]>=1)
    }
    else if (i<j) {
      viral_scores_two_comb[i,j] <- sum((viral_scores_two[,i]+viral_scores_two[,j])>=1)
    }
    else if (i>j) {
      viral_scores_two_comb[i,j] <- paste(round(sum((viral_scores_two[,i]+viral_scores_two[,j])>=1)/sum(viral_scores_two[,i]>=1 | viral_scores_two[,j]>=1)*100, digits=0), "%", sep="")
    }
  }
}

colnames(viral_scores_two_comb) <- c("VS2", "VS", "VB", "tv", "DVF", "tnv")
rownames(viral_scores_two_comb) <- c("VS2", "VS", "VB", "tv", "DVF", "tnv")
```

```{r}
viral_scores_two_comb
```

comparing VS2 and VIBRANT viruses - Figure S15
```{r}
vs2_vb <- data.frame(vs2=viral_scores_two[,1],
                     vb=viral_scores_two[,3],
                     comb_score=viral_scores_two[,1]+viral_scores_two[,3],
                     seqtype=viruses$seqtype,
                     max_score_group=viruses$max_score_group)

vs2_vb$comb[vs2_vb$comb_score<1] <- "neither"
vs2_vb$comb[vs2_vb$vs2==0.5 & vs2_vb$vb==0.5] <- "both_low"
vs2_vb$comb[vs2_vb$vs2>=1 & vs2_vb$vb==0.5] <- "vs2_high_vb_low"
vs2_vb$comb[vs2_vb$vs2>=0.5 & vs2_vb$vb==1] <- "vs2_low_vb_high"
vs2_vb$comb[vs2_vb$vs2>=1 & vs2_vb$vb==0] <- "vs2_only"
vs2_vb$comb[vs2_vb$vs2==0 & vs2_vb$vb>=1] <- "vb_only"
vs2_vb$comb[vs2_vb$vs2>=1 & vs2_vb$vb>=1] <- "both_high"

#vs2_vb$true_virus <- "not"
#vs2_vb$true_virus[vs2_vb$seqtype=="virus"] <- "virus"

vs2_vb_seqtype <- vs2_vb %>% 
  group_by(comb, comb_score, seqtype) %>%
  summarise(n = n())
```

```{r}
table(vs2_vb$comb)
```

```{r}
ggplot(vs2_vb_seqtype, aes(y=n, x=comb,
                   fill=as.factor(comb_score))) +
  geom_bar(stat="identity", color="black") +
  theme_light() +
  coord_flip() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    legend.position = "bottom"
  ) +
  scale_fill_brewer(palette = "PuOr", ) +
  xlab("") +
  ylab("Number of Sequences") + 
  facet_grid(~seqtype, scales = "free")
```

# overlap in proportion of contigs for tool sets - Figure S16

get number of tools and number of rules for each combination
```{r}
ann_col$numrules <- rowSums(ann_col[,1:6])
ann_col$numtools <- ann_col$numrules

for (i in 1:nrow(ann_col)) {
  if (ann_col$tv[i]==1) {
    ann_col$numtools[i] <- ann_col$numtools[i] + 3
  }

  if (ann_col$tnv[i]==1) {
    ann_col$numtools[i] <- ann_col$numtools[i] + 5
  }
}

ann_col$numtools[ann_col$numtools>6] <- 6

ann_col$tool <- rownames(ann_col)
```

```{r}
prop_high_viruses_long <- as.data.frame(prop_high_viruses)
prop_high_viruses_long$tool <- combos_list$toolcombo

prop_high_viruses_long <- as.data.frame(prop_high_viruses_long) %>% 
  pivot_longer(!tool,
               values_to="proportion",
               names_to="tool2") %>%
  inner_join(ann_col, by=c("tool"))

```{r}
head(prop_high_viruses_long)
```

```{r}
ggplot(prop_high_viruses_long[prop_high_viruses_long$vs2==1,], aes(x=as.factor(numrules), y=proportion, 
                                  color=as.factor(numrules), fill=as.factor(numrules))) +
  geom_boxplot() +
  #geom_point() +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Number of Rules") +
  ylab("Proportion of Contigs in Common") +
  scale_fill_manual(name="",
                     values = alpha(rev(viridis(6)), 0.3)) +
  scale_color_manual(name="",
                     values = alpha(rev(viridis(6)), 0.7))

ggplot(prop_high_viruses_long[prop_high_viruses_long$vs==1,], aes(x=as.factor(numrules), y=proportion, 
                                  color=as.factor(numrules), fill=as.factor(numrules))) +
  geom_boxplot() +
  #geom_point() +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Number of Rules") +
  ylab("Proportion of Contigs in Common") +
  scale_fill_manual(name="",
                     values = alpha(rev(viridis(6)), 0.3)) +
  scale_color_manual(name="",
                     values = alpha(rev(viridis(6)), 0.7))

ggplot(prop_high_viruses_long[prop_high_viruses_long$vb==1,], aes(x=as.factor(numrules), y=proportion, 
                                  color=as.factor(numrules), fill=as.factor(numrules))) +
  geom_boxplot() +
  #geom_point() +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Number of Rules") +
  ylab("Proportion of Contigs in Common") +
  scale_fill_manual(name="",
                     values = alpha(rev(viridis(6)), 0.3)) +
  scale_color_manual(name="",
                     values = alpha(rev(viridis(6)), 0.7))

ggplot(prop_high_viruses_long[prop_high_viruses_long$dvf==1,], aes(x=as.factor(numrules), y=proportion, 
                                  color=as.factor(numrules), fill=as.factor(numrules))) +
  geom_boxplot() +
  #geom_point() +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Number of Rules") +
  ylab("Proportion of Contigs in Common") +
  scale_fill_manual(name="",
                     values = alpha(rev(viridis(6)), 0.3)) +
  scale_color_manual(name="",
                     values = alpha(rev(viridis(6)), 0.7))
```
